---
title: "You Control This Rmd File"
author: "Cora Boyoung Jung, Enoch Mwesigwa, Jordan Severn"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
require(tidyverse)
require(ggformula)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(dev = c('pdf', 'png'),
                      fig.path = 'controlled-figures/legos-')
```

## Instructions

Alter each section to do as instructed.

Unless explicitly told to "remove this text" you can either remove, or leave in, the text instructions within each section. 

Leave the section headers undisturbed unless explicitly told to delete them.

## Legos Data

In this section, read in the legos dataset from <https://sldr.netlify.app/data/legos.csv> **but make sure that the code you use, and any messages it generates, do not show up in the output file.**

```{r, echo = FALSE, message = FALSE}
legos <- read_csv('https://sldr.netlify.app/data/legos.csv') %>%
  mutate(Theme = factor(Theme),
         Subtheme = factor(Subtheme))
```

## A Mal-Sized Figure

Make one figure (any figure of your choice) using the legos dataset.

Ensure that the code used to generate is **is shown** in the knitted file.

Also ensure **that the figure itself is a bit too small** and **the font size used in the figure is too big, especially relative to the figure size.**

```{r too-small, fig.height = 1, fig.width = 5}
theme_set(theme_minimal(base_size = 30))
gf_point(Price ~ Minifigures, 
         alpha = 0.1,
         data = legos)
```


## A Better Proportioned Figure

Make the same figure again.

But this time, make it a **reasonable size** and with a **reasonable-size** font.

And this time, make sure the **code used to generate it is not shown** in the knitted file.

```{r better-sized, fig.height = 3, fig.width = 5,  echo = FALSE}
theme_set(theme_minimal(base_size = 11))
gf_point(Price ~ Minifigures, 
         alpha = 0.1,
         data = legos)
```

## This is Silly

Add code to create another figure in this section (your choice, but different from what you have done so far) with the lego data.

But make sure **the code to make the figure is shown in the knitted file** (just shown, without commenting it out) and **the figure itself is not shown.**

```{r silly, fig.show = 'hide'}
gf_boxplot(Price ~ factor(Year), 
         alpha = 0.1,
         data = legos) %>%
  gf_theme(axis.text.x = element_text(angle = 70))
```


## Saving Those Figs

Revise your file (you'll make changes to your settings chunk) so that PDF and jpeg image files of each figure get saved to a folder called "controlled-figures" with file name prefix "legos-" (followed by the R code chunk label).

Make sure the R code chunk labels (and thus the figure file names) include "too-small", "better-sized", and "silly" respectively. (You are encouraged to make the labels more informative too, like if you made a histogram of number of pieces you might label that chunk "too-small-npieces-hist".)

Make sure to examine your Files tab to verify where the figures get saved and how the names work.

## Ta-Da!

You did it. You now control this file.

If you have more time left after handing in your un-knitted Rmd file on Moodle, you may use if for project work and coordination.