---
title: "UMCH"
author: "Cora Boyoung Jung, Enoch Mwesigwa, Jordan Severn"
output:
  pdf_document: default
  html_document:
    df_print: paged
---


```{r setup, include=FALSE}
library(plyr)  ## or dplyr (transform -> mutate) 
require(mosaic)
require(ggformula)
require(tidyverse)
require(s245)
require(DHARMa)
library(reticulate) # for using python
#install.packages('Rmisc', dependencies = TRUE) #good summary tool
require(Rmisc)

knitr::opts_chunk$set(echo = TRUE, 
                      error = TRUE,
                      message = FALSE,
                      fig.width = 10, fig.height = 6)
knitr::opts_chunk$set(dev = c('png'),
                      fig.path = 'UMCH-figures/')

theme_set(theme_minimal(base_size = 12))
```

## Import csv
  
```{r}
UMCH.raw <- read_csv("https://raw.githubusercontent.com/Cora-Boyoung-Jung/UMCH/main/data/UMCH_p.csv",
                     col_types = cols(Birthdate = col_date(format = "%m/%d/%Y"),
                                      Age = col_integer()))
glimpse(UMCH.raw)
```
## Tidying data
  
```{r}
UMCH.clean <- UMCH.raw %>% mutate(Status = tolower(Status))

v1 <- unlist(UMCH.clean$AgeGroup)
UMCH.clean$AgeGroup <- relist(replace(v1, v1=="two_year", "2yo"), skeleton=UMCH.clean$AgeGroup)

v1 <- unlist(UMCH.clean$AgeGroup)
UMCH.clean$AgeGroup <- relist(replace(v1, v1=="three_year", "3yo"), skeleton=UMCH.clean$AgeGroup)

v1 <- unlist(UMCH.clean$AgeGroup)
UMCH.clean$AgeGroup <- relist(replace(v1, v1=="four_year", "4yo"), skeleton=UMCH.clean$AgeGroup)

neworder <- c("infant","toddler","2yo", "3yo", "4yo")

UMCH.clean <- arrange(transform(UMCH.clean,
             AgeGroup=factor(AgeGroup,levels=neworder)),AgeGroup)

names(UMCH.clean)[13] <- "PhysDev %"
names(UMCH.clean)[14] <- "LangDev %"
names(UMCH.clean)[15] <- "AdaptiveSelfHelp %"
names(UMCH.clean)[16] <- "AdaptiveSocioEmotional %"
names(UMCH.clean)[17] <- "AcademicCognitive %"
names(UMCH.clean)[18] <- "AcademicCognitiveMath %"
names(UMCH.clean)[19] <- "AcademicCognitiveLit %"
```

## Exploring data set
  
```{r}
head(UMCH.clean)
summary(UMCH.clean)
```

## Graphics

### Which domain in the areas of development is scored the lowest and highest in which age group and overall?  

```{r}
UMCH.long <- UMCH.clean %>%
  subset(select = -c(Filename, Birthdate, Age, Gender, TotalScore, Status, Examiner,Adaptive_SelfHelp, Adaptive_SocialEmotional, LanguageDevelopment,AcademicAndCognitive,PhysicalDevelopment, AcademicAndCognitive_Literacy, AcademicAndCognitive_Maths)) %>%
  pivot_longer(!AgeGroup, names_to = "Domain", values_to = "Score") %>%
  na.omit() %>%
  mutate(AgeGroup = as.factor(AgeGroup),
         Domain = as.factor(Domain),
         Score = as.numeric(Score))

UMCH.stat <- UMCH.long %>%
  dplyr::group_by(AgeGroup,Domain) %>%
  dplyr::summarise(Min=min(Score),
                   Max=max(Score),
                   Mean=round(mean(Score), digits = 2),
                   Median=round(median(Score), digits = 2))

UMCH.stat %>%
  knitr::kable()
```
```{r totalScore_by_AgeGroup_and_Domain}
UMCH.between <- summarySE(data=UMCH.long, measurevar="Score", groupvars=c("Domain", "AgeGroup"), na.rm=FALSE, conf.interval=.95)

ggplot(UMCH.between, aes(x = Score,  y = Domain, color=AgeGroup)) +
  geom_point() + 
  geom_errorbar(width=.1, aes(xmin=Score-ci, xmax=Score+ci), colour='#00BFC4') +
  facet_wrap(~AgeGroup) +
  labs(title="Percent of total Score of Children Grouped by AgeGroup and Domain", y="Score (%)") +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1)) +
  expand_limits(y = 0) 
```


### How does score vary by age group?  

```{r totalscore-by-agegroup}
gf_boxplot(~TotalScore, data=UMCH.clean, fill='#00BFC4', color='black') %>% 
   gf_jitter(color = 'grey', alpha = 1, width= 0.15) %>%
  gf_labs(title="Total Scores by Age Group", x="Total Score", y="Counts") + facet_wrap(~AgeGroup) 
  
```

### How does score vary by gender?  

```{r totalscore-by-gender-and-agegroup}
ggplot(UMCH.clean, aes(x = Gender,  y = TotalScore, fill=Gender)) %>%
  gf_jitter(color = 'grey', alpha = 1, width= 0.15)
  geom_boxplot() + 
  facet_wrap(~AgeGroup) +
  labs(title="Total Scores by Gender for Each AgeGroup", x="Gender") +
  expand_limits(y = 0) 
  
```


## TotalScore Model

We used linear regression model because it is used to show/predict the relationship between variables where the response variable is continuous. We can use up to 2 predictors given the size of the data (36/15). We will not add any interaction or random effect due to the size of the dataset. Our predictors will be Age and Gender because all other variables contributes to the Total Score which does not make logical sense to include those.  

Response variable: TotalScore  
Predictor(s): Age, Gender   
Regression model: Linear regression 

```{r model plan}
mod_total <- lm(TotalScore ~ Age + Gender,  
          data = UMCH.clean)  
summary(mod_total)  
```

## Model Assessment

```{r linearity-and-error-variance}
gf_point(resid(mod_total) ~ fitted(mod_total)) %>%
  gf_labs(x = 'Fitted Values', y = 'Residuals')
```

The fitted values vs. the residual plot is used for linearity and error variance. There is no trumpeting present so it passes the error variance condition. The vertical lines are not a problem, because it is the result of the predictor, age, is rounded up to the nearest year, so it shows like the predictor is categorical. So, it passes both the linearity and error variance conditions.  

```{r independence}
acf(resid(mod_total))
```

Looking at the acf plot, the plot passes the independence condition. All the lines are within the dotted lines and the lag at 0 is normal. 

```{r normality}
gf_histogram(~resid(mod_total), bins = 5) %>%
  gf_labs(x = "Residuals")
```

Since the data is limited, the bins need to be limited. The histogram passes the normality condition. 


## Prediction Plot

```{r predictions}
# make hypothetical data
gender_pred_data <- expand.grid(Age = seq(from = 0, by = 1, to = 4), 
                                Gender = pull(UMCH.clean, Gender) %>% factor() %>% levels())

# make predictions with SEs
preds <- predict(mod_total,
                 type = 'response',
                 se.fit = TRUE,
                 newdata = gender_pred_data)

# put predictions and CI bounds in data set
gender_pred_data <- gender_pred_data %>%
  mutate(fitted = preds$fit,
         CI_low = fitted - 1.96 * preds$se.fit,
         CI_high = fitted + 1.96 * preds$se.fit)

# make a graph
gf_line(fitted ~ Age,
         color = ~Gender,
         data = gender_pred_data) %>%
  gf_labs(x = 'Age', y = 'Predicted Total Score') %>%
  gf_ribbon(CI_low + CI_high ~ Age,
            fill = ~Gender,
            color = ~Gender)
```
According to the prediction plot, there seems to be a positive trend between the score and age where as age increases, the score also increases. This makes sense because the "pass" score for each age group differs according to the age with lower age having the lowest pass score. Also, it looks like the total score is a little higher for female compared to male. However, the overlapping range is high and the number of children is not the same (21 females and 15 males), so there is some extend of uncertainty.  






